% There is currently a problem with spacing somewhere so that Table of
% Contents, List of Tables, and List of Figures have the wrong amount
% of space.  Others are OK though...
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

Luke Harmon has been a tremendous mentor, teacher, collaborator, and friend. He came to know me better than I knew myself. He found a way to bring out the best in me and restrain the worst. And he made science fun---there is no one else I would rather make mistakes with. 

I thank my committee: Jack Sullivan, for supporting and challenging me in science and in life; Scott Nuismer, for making sure I knew where I stood; Paul Joyce, for always making time when I got stuck on a problem; and Arne Mooers, for introducing me to evolutionary biology and for keeping me honest over the years.

David Tank was practically a second advisor to me. He took me under his wing and taught me how to think and how to be a scientist. And somehow he was able to convince me that plants are actually pretty damn awesome (I will never forget digging up fossils in Clarkia). I am also grateful to Larry Forney, for taking the time to talk to me, and for calling me on my bullshit when I needed to be called on it.

Simon Uribe-Convers, Brice Sarver, Travis Hagey, and Tyler Hether were my brothers in the trenches---we worked late together, argued endlessly about biology and everything else, and kept each other sane during the ups and downs of grad school. I am extremely grateful for their friendship.

There is an old and persistent (not to mention, pernicious) myth that working towards a doctorate is, or at least ought to be, a lonely journey. My own experience could not be farther from this; collaboration and camaraderie have been central to all aspects of my intellectual life. A great deal of this work, and so much of the thought behind it, has sprung from my interactions with other (and, in many cases, young) scientists. In particular, I would like to thank Rich FitzJohn and Josef Uyeda. They challenged me to think harder and to never take the easy way out. I have learned so much from both of them and enjoyed ever minute (or at least, most) of the many that we spent together brainstorming ideas, writing papers, and hacking code. I also thank Jon Eastman for getting off his rocket ship to help me whenever I fell off my tricycle. And David Bapst has been my constant sounding board and sparring partner throughout my dissertation. It is only because of him that I can converse with paleontologists without complete embarrassment. I would also specifically like to thank some of my collaborators and mentors---Michael Alfaro, Steve Arnold, Frank Burbrink, Will Cornwell, Bernie Crespi, Joe Felsenstein, David Green, Paul Harnik, Mark Kirkpatrick, Craig Miller, Brian O'Meara, Sally Otto, Erica Bree Rosenblum, Carl Simpson, Graham Slater, David Swofford, and Amy Zanne---whose intellectual and personality generosity is deeply humbling. 

It has been an incredible privilege, and a whole lot of fun, to do science in the Harmon/RoHa lab. You guys are the best: I thank Joseph Brown, Daniel Caetano, Simone Des Roches, Kayla Hardwick, Denim Jochimsen, Suzanne Joneson, Rafael Maia, Eliot Miller, Tom Poorten, James Rosindell, and Jamie Voyles.

I loved being an IBESTian at the University of Idaho. I am thankful to all the people who made it so fun, both in and out of the office. In particular, I would like to thank Daniel Beck, Lucius Caldwell, Matthieu Delcourt, Anahi Espindola, Michael France, Sam Hunter, Sarah Jacobs, CJ Jenkins, Maribeth Latvis, Wesley Loftie-Eaton, Hannah Marx, Genevieve Metzger, Diego Morales Briones, Matt Settles, Matt Singer, Chloe Stenkamp-Strahm, and ET Thornquist. I also want to give an extra shoutout to all the members, past and present, of the Phylogenetics Reading Group (PuRGe); I loved arguing with you all for the past five years. I thank all the IBEST staff and faculty, especially Lisha Abendroth and Eva Top, for without their boundless patience, I would literally not have finished my Ph.D.

I thank Roxana Hickey for all the things: for celebrating with me when things were going well; for commiserating when things were not; for listening to me rant even when she didn't care; for being my friend; for suffering me when I was insufferable; and for always pushing me to be better. Last, I would like to thank my family: my parents, Richard and Brenna Pennell, and my siblings, Stephen, Alec, and Laura. I cannot even begin to articulate the support they have given me over the years nor my gratitude for it. For once, I have no words.


%%% Local Variables:
%%% TeX-master: "thesis.tex"
%%% TeX-PDF-mode: t
%%% End:
